---
title: "Foundations of R Programming for Data Science"
#subtitle: "Data Science"
#author:
#   - name: Meghan Hall
#   - url: https://meghan.rbind.io
#   - affiliation: Carnegie Mellon University
#   - date: "`r Sys.Date()`"
description: |
   This is a comprehensive training program designed to equip participants with the essential skills and knowledge required to effectively use the R programming language for data science applications from different domains.
site: distill::distill_website
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Overview

This course covers fundamental concepts, practical techniques, and hands-on exercises to ensure that attendees gain a solid foundation in R programming and its applications in data analysis, visualization, and statistical modeling.

## Intention and Motivation

The primary intention of this training is to introduce participants to the R programming language and provide them with the necessary tools to perform data science tasks. By the end of the training, participants will be able to write R code, manipulate data, create visualizations, and apply statistical techniques to real-world datasets.


## Objectives

- **Introduction to R**: Familiarize participants with the R programming environment, including installation, basic syntax, and key features.
- **Data Manipulation**: Teach participants how to import, clean, and manipulate data using R's powerful data manipulation packages such as dplyr and tidyr.
- **Data Visualization**: Enable participants to create informative and visually appealing data visualizations using ggplot2 and other visualization libraries.
- **Statistical Analysis**: Provide a foundational understanding of statistical concepts and how to apply them using R, including descriptive statistics, hypothesis testing, and regression analysis.
- **Reproducible Research**: Introduce the concept of reproducible research using R Markdown, enabling participants to create dynamic and shareable reports.
- **Practical Applications**: Engage participants in hands-on exercises and projects to apply their learning to real-world scenarios and datasets.


## Lectures

Lecture slides are available below in html format, and the .Rmd files are linked from GitHub.

| Days | Topic                                          | html | Rmd  |
|---------|------------------------------------------------|------|------|
| Day0  | Welcome and Training Overview| [`r icons::fontawesome$solid$laptop`]((https://mgisa.github.io/RTraining/trial/Days_displays#1) | [`r icons::fontawesome$brands$github`](https://github.com/mgisa/RTraining/blob/main/slides/Days_displays.qmd) |
| Day2  | the grammar of graphics  | [`r icons::fontawesome$solid$laptop`](https://cmu-36315.netlify.app/slides/36-315_summer2021_lec2#1) | [`r icons::fontawesome$brands$github`](https://github.com/meghall06/CMU-36-315-site/blob/main/slides/36-315_Summer2021_Lec2.Rmd)|
| Day3  | bar graphs | [`r icons::fontawesome$solid$laptop`](https://cmu-36315.netlify.app/slides/36-315_summer2021_lec3#1) |  [`r icons::fontawesome$brands$github`](https://github.com/meghall06/CMU-36-315-site/blob/main/slides/36-315_Summer2021_Lec3.Rmd)|
| Day4  | visualizing distributions | [`r icons::fontawesome$solid$laptop`](https://cmu-36315.netlify.app/slides/36-315_summer2021_lec4#1) |  [`r icons::fontawesome$brands$github`](https://github.com/meghall06/CMU-36-315-site/blob/main/slides/36-315_Summer2021_Lec4.Rmd)|
| Day5  | line graphs & working with time | [`r icons::fontawesome$solid$laptop`](https://cmu-36315.netlify.app/slides/36-315_summer2021_lec5#1) |  [`r icons::fontawesome$brands$github`](https://github.com/meghall06/CMU-36-315-site/blob/main/slides/36-315_Summer2021_Lec5.Rmd)|
| 6  | scatter plots & relational data | [`r icons::fontawesome$solid$laptop`](https://cmu-36315.netlify.app/slides/36-315_summer2021_lec6#1) |  [`r icons::fontawesome$brands$github`](https://github.com/meghall06/CMU-36-315-site/blob/main/slides/36-315_Summer2021_Lec6.Rmd)|
| 7  | grab bag (pie charts, heat maps, maps) | [`r icons::fontawesome$solid$laptop`](https://cmu-36315.netlify.app/slides/36-315_summer2021_lec7#1) |  [`r icons::fontawesome$brands$github`](https://github.com/meghall06/CMU-36-315-site/blob/main/slides/36-315_Summer2021_Lec7.Rmd)|
| 8  | taking plots to the next level | [`r icons::fontawesome$solid$laptop`](https://cmu-36315.netlify.app/slides/36-315_summer2021_lec8#1) |  [`r icons::fontawesome$brands$github`](https://github.com/meghall06/CMU-36-315-site/blob/main/slides/36-315_Summer2021_Lec8.Rmd)|
| 9  | colors, fonts, annotations, themes | [`r icons::fontawesome$solid$laptop`](https://cmu-36315.netlify.app/slides/36-315_summer2021_lec9#1) |  [`r icons::fontawesome$brands$github`](https://github.com/meghall06/CMU-36-315-site/blob/main/slides/36-315_Summer2021_Lec9.Rmd)|
| 10  | extensions to `ggplot2` | [`r icons::fontawesome$solid$laptop`](https://cmu-36315.netlify.app/slides/36-315_summer2021_lec10#1) |  [`r icons::fontawesome$brands$github`](https://github.com/meghall06/CMU-36-315-site/blob/main/slides/36-315_Summer2021_Lec10.Rmd)|
| 11  | midterm review | [`r icons::fontawesome$solid$laptop`](https://cmu-36315.netlify.app/slides/36-315_summer2021_lec11#1) |  [`r icons::fontawesome$brands$github`](https://github.com/meghall06/CMU-36-315-site/blob/main/slides/36-315_Summer2021_Lec11.Rmd)|
| 12  | model output visualizations from `tidymodels` & user-defined functions | [`r icons::fontawesome$solid$laptop`](https://cmu-36315.netlify.app/slides/36-315_summer2021_lec12#1) |  [`r icons::fontawesome$brands$github`](https://github.com/meghall06/CMU-36-315-site/blob/main/slides/36-315_Summer2021_Lec12.Rmd)|
| 13  | text analysis with `tidytext` and tables with `kableExtra` | [`r icons::fontawesome$solid$laptop`](https://cmu-36315.netlify.app/slides/36-315_summer2021_lec13#1) |  [`r icons::fontawesome$brands$github`](https://github.com/meghall06/CMU-36-315-site/blob/main/slides/36-315_Summer2021_Lec13.Rmd)|
| 13  | tables with `gt` and `gtsummary` | [`r icons::fontawesome$solid$laptop`](https://cmu-36315.netlify.app/slides/36-315_summer2021_lec13_addl) |  [`r icons::fontawesome$brands$github`](https://github.com/meghall06/CMU-36-315-site/blob/main/slides/36-315_Summer2021_Lec13_addl.Rmd)|
| 14  | presentations with `xaringan` | [`r icons::fontawesome$solid$laptop`](https://cmu-36315.netlify.app/slides/36-315_summer2021_lec14#1) |  [`r icons::fontawesome$brands$github`](https://github.com/meghall06/CMU-36-315-site/blob/main/slides/36-315_Summer2021_Lec14.Rmd)|
| 15  | animation with `gganimate`, interactivity with `plotly`, and iteration with `purrr` | [`r icons::fontawesome$solid$laptop`](https://cmu-36315.netlify.app/slides/36-315_summer2021_lec15#1) |  [`r icons::fontawesome$brands$github`](https://github.com/meghall06/CMU-36-315-site/blob/main/slides/36-315_Summer2021_Lec15.Rmd)|

## Labs

Lab slides are available below in html format, and the .Rmd files are available at the GitHub repo linked in the nav bar above.

- [Lab 1](slides/36-315_Summer2021_Lab1_lec.html)
- [Lab 2](slides/36-315_Summer2021_Lab2_lec.html)
- [Lab 3](slides/36-315_Summer2021_Lab3_lec.html)
- [Lab 4](slides/36-315_Summer2021_Lab4_lec.html)
- [Lab 5](slides/36-315_Summer2021_Lab5_lec.html)
- [Lab 6](slides/36-315_Summer2021_Lab6_lec.html)
- [Lab 8](slides/36-315_Summer2021_Lab8_lec.html)
- [Lab 9](trial/NEAIR.html)

## Resources

### Data Visualization

+ Fundamentals of Data Visualization: <https://clauswilke.com/dataviz/index.html>

+ Data Visualization: A Practical Introduction: <https://socviz.co/index.html#preface>

### The tidyverse

+ Best resource for overall analysis in R using the tidyverse: <https://r4ds.had.co.nz/>

### `ggplot2`

+ `ggplot2` book: <https://ggplot2-book.org/>

+ The `ggplot2` cheatsheet: <https://www.maths.usyd.edu.au/u/UG/SM/STAT3022/r/current/Misc/data-visualization-2.1.pdf>

+ Details of various charts made in R: <https://www.r-graph-gallery.com/index.html>

+ Creating custom themes: <https://themockup.blog/posts/2020-12-26-creating-and-using-custom-ggplot2-themes/>

+ R Graphics cookbook: <https://r-graphics.org/>

+ A guide to various scales: <https://ggplot2tor.com/scales>

### R Markdown

+ The definitive guide to R Markdown: <https://bookdown.org/yihui/rmarkdown/>

+ R Markdown reference guide: <https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf>

+ R Markdown cheat sheet: <https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf>


## Acknowledgments {.appendix}

This website is made with the [distill package](https://rstudio.github.io/distill/). Many thanks to [Alison Hill](https://alison.rbind.io/) and [Desirée De Leon](https://desiree.rbind.io/) and their [Teaching in Production](https://rstudio-education.github.io/teaching-in-production/) website, which provided the inspiration for this site. Also thanks to Tom Mock for his enormously helpful [introduction to distill](https://themockup.blog/posts/2020-08-01-building-a-blog-with-distill/).

```{r, include=FALSE, results='asis'}
library(metathis)
meta() %>%
  meta_social(
    title = "CMU 36-315",
    description = "Course website for Statistical Graphics & Visualization, by Meghan Hall",
    url = "https://cmu-36315.netlify.app/",
    image = "https://raw.githubusercontent.com/meghall06/CMU-36-315-site/main/icon-192.png",
    image_alt = "CMU 36-315",
    og_type = "website",
    og_author = "Meghan Hall",
    twitter_card_type = "summary",
    twitter_creator = "@meghanmhall"
  )
```