---
title: "Day1"
subtitle: ""  
author: "Murera Gisa"
date: 
output:
  xaringan::moon_reader:
    seal: false
    css: ["xaringan-themer.css", "custom.css"]
    lib_dir: libs
    nature:
      titleSlideClass: ["left", "middle"]
      highlightStyle: github
      highlightLines: true
      highlightSpans: true
      countIncrementalSlides: false
      slideNumberFormat: "%current%"
---
```{r setup, include=FALSE}
#library(emo)
#library(icons)
library(kableExtra)
#library(betweenthepipes)
library(tidyverse)
library(showtext)
options(knitr.kable.NA = '')
options(htmltools.dir.version = FALSE, htmltools.preserve.raw = FALSE)
knitr::opts_chunk$set(fig.retina = 3, 
                      warning = FALSE, 
                      message = FALSE,
                      fig.path = "figs/Lec1/")

font_add_google(name = "Open Sans",   # Name of the font on the Google Fonts site
                family = "open_sans")
font_add_google(name = "Source Serif Pro",   # Name of the font on the Google Fonts site
                family = "source_serif_pro")
showtext_auto()


cmu_theme <- function () { 
  theme_linedraw(base_size=11, base_family="source_serif_pro") %+replace% 
    theme(
      panel.background  = element_blank(),
      plot.background = element_rect(fill = "transparent", color = NA), 
      legend.background = element_rect(fill = "transparent", color = NA),
      legend.key = element_rect(fill = "transparent", color = NA),
      axis.ticks = element_blank(),
      panel.grid.major = element_line(color = "grey90", size = 0.3), 
      panel.grid.minor = element_blank(),
      plot.title = element_text(size = 18, hjust = 0, vjust = 0.5, face = "bold", margin = margin(b = 0.2, unit = "cm")),
      plot.subtitle = element_text(size = 12, hjust = 0, vjust = 0.5, margin = margin(b = 0.2, unit = "cm")),
      plot.caption = element_text(size = 7, hjust = 1, face = "italic", margin = margin(t = 0.1, unit = "cm")),
      axis.text.x = element_text(size = 13),
      axis.text.y = element_text(size = 13)
    )
}

```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)

style_mono_accent(
  header_font_google = google_font("Open Sans"),
  text_font_google   = google_font("Source Serif Pro", "400", "400i"),
  code_font_google   = google_font("Roboto Mono"),
  link_color = "#bb0",
  base_color = "#b9770e",
  extra_fonts = list(
    google_font("Lato")
  ),
  extra_css = list(
    ".title-slide h2" = list("font-family" = "Source Serif Pro"),
    ".title-slide h3" = list("font-family" = "Source Serif Pro")
  )
)
```

class: inverse, center, middle

##  Welcoming and Housekeeping


Trainer: Murera Gisa <br> Senior Data Scientist <br> The National Bank of Rwanda, BNR <br> September, 2024

---

## Training Team and Time
<br>
<br>

-  Trainer: `Murera Gisa`
<br>

- Logistics: `Hillary Kiplagat`
<br>

- Training hours: 9:00AM-4:30PM
<br>

---

## Welcome Note
<br>
<br>

- Hillary brings on board the trainees, addresses the facilities, and Logistics.

--

- Together we can highlight the room golden rules for effective training.  

--

- Together we can also highlight the key expectations.

---
class: center, middle

## Training Tools

<br>
<br>

.large[`R` and  `RStudio`]

---
class: center, middle

## Requirements


Personal Computer
 
--

Stable Internet

--

Eager and Motivation to explore new skills and Knowledge

--

---
class: center, middle

## Material

Slides and Collaborative Discussions

--

Labs and Exercises

--

Assignments

--

Group Project and Presentation

---

class: inverse, center, middle

## Training Components

```{r component-graph, echo = FALSE, fig.align = "center", fig.height = 6.5}
course_components <- tribble(
  ~Component, ~Percent,
  "Slides", 0.4,
  "Assignments", 0.1,
  "Labs and Exercises", 0.3,
  "Final Group Project Report", 0.15,
  "Final Group Presentation", 0.05
)

course_components %>% 
  mutate(component = str_replace(Component, " ", "\n")) %>% 
  ggplot(aes(x = reorder(component, -Percent), y = Percent)) +
  geom_bar(stat = "identity", fill = "#ba4") +
  labs(title = "Weight of training components",
       subtitle = "Summer School 2024",
       caption = "#Code4Cause") +
  geom_text(aes(reorder(component, -Percent), Percent, label = scales::percent(Percent, accuracy = 1)), 
            family = "source_serif_pro", fontface = 2, vjust = -0.75) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0, 0.52)) +
  cmu_theme() +
  theme(axis.title = element_blank(),
        panel.grid.major.x = element_blank(),
        axis.text.x = element_text(size = 11))
```

---

## Other Logistics 

<br>
<br>

[Training Website(s)](mgisa.github.io/RTraining/) that hosts:

--

- Training overviews, objections and its relevance

--

- Training ground rules

--

- Who is Trainer?

--

- Training materials

---

## Upcoming Training


- Database and R programming

--

- R package development for data management

--

- Dashboarding and deployment ( Embedding `PowerBI` into `R` and the way around)

--

- NLP, Sentiment Analysis and Tone of Communication

--

- Social Media Network Interaction

--

- Machine Learning: Development

--
- Prediction, Forecasting and Deployment (MLOps)

--

- Upon to Client Request

---

class: inverse, center, middle

.Large[ 🎯 Thank you, and let's get our hands dirty with R coding!!! ]